---
title: "DORSAL SCARS vs Approximate Area and MODULE"
author: David NOra
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load Libraries
```{r}
library(car)
library(dplyr)
library(broom)
library(ggplot2)
library(readr)
```

## Load the Data
```{r load-data}
# Load and label the datasets
kalavan_t1 <- read_csv("C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\Theory and Methods\\Submission\\Moduleflakescategories\\analysis\\data\\raw_data\\kalavan_2_T1_Edge_Modification.csv")

kalavan_t2 <- read_csv("C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\Theory and Methods\\Submission\\Moduleflakescategories\\analysis\\data\\raw_data\\Kalavan_2_T2_Edge_Modification.csv")

ararat_1 <- read_csv("C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\Theory and Methods\\Submission\\Moduleflakescategories\\analysis\\data\\raw_data\\Ararat_1_Cave_Edge_Modification.csv") 

```

## Prepare Data
```{r prepare-data}
# Ensure the grouping variable is a factor
kalavan_t1$MODULE <- as.factor(kalavan_t1$MODULE)
kalavan_t2$MODULE <- as.factor(kalavan_t2$MODULE)
ararat_1$MODULE <- as.factor(ararat_1$MODULE)

# Check structure
str(kalavan_t1)
str(kalavan_t2)
str(ararat_1)

kalavan_t1_complete <- kalavan_t1 %>%
  filter(!is.na(`DORSAL SCARS`) & !is.na(`Approximate area`) & !is.na(MODULE))
kalavan_t2_complete <- kalavan_t2 %>%
  filter(!is.na(`DORSAL SCARS`) & !is.na(`Approximate area`) & !is.na(MODULE))
ararat_1_complete <- ararat_1 %>%
  filter(!is.na(`DORSAL SCARS`) & !is.na(`Approximate area`) & !is.na(MODULE))
```

## MANOVA: Multivariate Test of Group Differences
```{r manova-test}
# Run the MANOVA for each dataset
manova_model1 <- manova(cbind(`DORSAL SCARS`, `Approximate area`) ~ MODULE, data = kalavan_t1_complete)
manova_model2 <- manova(cbind(`DORSAL SCARS`, `Approximate area`) ~ MODULE, data = kalavan_t2_complete)
manova_model3 <- manova(cbind(`DORSAL SCARS`, `Approximate area`) ~ MODULE, data = ararat_1_complete)

# Get Wilks' Lambda summaries
wilks_summary1 <- summary(manova_model1, test = "Wilks")
wilks_summary2 <- summary(manova_model2, test = "Wilks")
wilks_summary3 <- summary(manova_model3, test = "Wilks")

wilks_summary1
wilks_summary2
wilks_summary3
```

## Export Wilks' Lambda Test to CSV (Manual Extraction)
```{r export-wilks}
# Extract and save Wilks' statistics for each dataset
extract_wilks <- function(summary_result) {
  data.frame(
    Test = "Wilks",
    ApproxF = summary_result$stats[1, "approx F"],
    NumDF = summary_result$stats[1, "num Df"],
    DenDF = summary_result$stats[1, "den Df"],
    PrF = summary_result$stats[1, "Pr(>F)"]
  )
}

write.csv(extract_wilks(wilks_summary1), "C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\Theory and Methods\\Submission\\Moduleflakescategories\\analysis\\data\\derived_data\\wilks_lambda_results_kalavan_t1.csv", row.names = FALSE)
write.csv(extract_wilks(wilks_summary2), "C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\Theory and Methods\\Submission\\Moduleflakescategories\\analysis\\data\\derived_data\\wilks_lambda_results_kalavan_t2.csv", row.names = FALSE)
write.csv(extract_wilks(wilks_summary3), "C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\Theory and Methods\\Submission\\Moduleflakescategories\\analysis\\data\\derived_data\\wilks_lambda_results_ararat_1.csv", row.names = FALSE)
```

## Post-hoc Analyses: Univariate ANOVAs and Tukey Tests

Once the MANOVA shows a significant multivariate effect, we follow up with univariate ANOVAs and post-hoc tests to explore which variables and group comparisons contribute most to the effect.

### ANOVA and Tukey HSD for `DORSAL SCARS`

```{r anova-scars}
# Kalavan T1
aov_scars1 <- aov(`DORSAL SCARS` ~ MODULE, data = kalavan_t1_complete)
summary(aov_scars1)
tukey_scars1 <- TukeyHSD(aov_scars1)
tukey_scars1

# Kalavan T2
aov_scars2 <- aov(`DORSAL SCARS` ~ MODULE, data = kalavan_t2_complete)
summary(aov_scars2)
tukey_scars2 <- TukeyHSD(aov_scars2)
tukey_scars2

# Ararat 1
aov_scars3 <- aov(`DORSAL SCARS` ~ MODULE, data = ararat_1_complete)
summary(aov_scars3)
tukey_scars3 <- TukeyHSD(aov_scars3)
tukey_scars3
```

### ANOVA and Tukey HSD for `Approximate area`

```{r anova-area}
# Kalavan T1
aov_area1 <- aov(`Approximate area` ~ MODULE, data = kalavan_t1_complete)
summary(aov_area1)
tukey_area1 <- TukeyHSD(aov_area1)
tukey_area1

# Kalavan T2
aov_area2 <- aov(`Approximate area` ~ MODULE, data = kalavan_t2_complete)
summary(aov_area2)
tukey_area2 <- TukeyHSD(aov_area2)
tukey_area2

# Ararat 1
aov_area3 <- aov(`Approximate area` ~ MODULE, data = ararat_1_complete)
summary(aov_area3)
tukey_area3 <- TukeyHSD(aov_area3)
tukey_area3
```
